---
title: "First Assignment"
author: "Nishanth Ravindran"
date: "25 July 2016"
output:
  html_document: default
  pdf_document:
    fig_caption: yes
    fig_crop: no
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# =====================================================
# R: HW 1
# Description: Basics of character vectors
# Data: Teams that played the superbowl
# =====================================================

# Write your full name
# Name: Nishanth Ravindran
# Start a new R session (in RStudio)

#Load the given data using load()

```{r}
load("~/R Working Directory/superbowl_teams.RData")
```

# inspect the objects with ls()
# (you should be able to see three vectors)

```{r}
ls()
```

# year: year of superbowl from 1967 to 2015
# winner: champions
# loser: losing teams

# find the class of each vector (year, winner, loser)
```{r}
class(year)
class(winner)
class(loser)
```
# use length() to determine whether the three vectors
# have the same number of elements
```{r}
length(year)
length(winner)
length(loser)
if(length(year)==length(loser)&&length(year)==length(winner)){print("The Length is same")}
```
# =====================================================
# Winning teams
# Write the commands to answer the following questions
# =====================================================

# get the first 5 champions
```{r}
head(winner)
```

# get the last 5 champions
```{r}
tail(winner)
```

# how many unique champions?
```{r}
unique(winner)
```
# sort unique champions alphabetically (from A to Z)
```{r}
sort(unique(winner),decreasing = FALSE)
```


# decreasingly sort unique champions alphabetically (from Z to A)
```{r}
sort(unique(winner),decreasing = TRUE)

```


# get the champions in even positions (2, 4, 6, 8, etc)
```{r}
winner[seq_along(winner) %%2 == 0]
```


# get the champions in odd positions (1, 3, 5, 7, etc)
```{r}
winner[seq_along(winner) %%2 > 0]
```

# use the function table() to get a 
# table of frequencies for the winning teams
# (assign the table to the object 'win_freqs')
```{r}
win_freqs<-table(winner)
win_freqs<-sort(win_freqs,decreasing = TRUE)
win_freqs
```


# what team has won the superbowl most times?
# and how many times?

```{r}
win_freqs[1]
```


# apply table() on 'win_freqs', this will give you
# how many teams have won how many superbowls
```{r}
win_freqs
```



# create a vector 'champions' by sorting the frequencies
# 'win_freqs' in decreasing order
```{r}
win_freqs<-sort(win_freqs,decreasing = TRUE)
champions<-as.vector(win_freqs)
champions
```


# make a barplot of 'champions' with barplot()
```{r}
barplot(champions,width=2,space=0.5)
```


# What are the championships (i.e. years) of "San Francisco 49ers"
```{r}
winner_year<-data.frame(winner,year)

print(subset(winner_year,winner_year$winner=="San Francisco 49ers"))

```


# What are the championships (i.e. years) of "Oakland Raiders" 
```{r}
print(subset(winner_year,winner_year$winner=="Oakland Raiders"))
```


# when was the last time Denver Broncos won a superbowl?
```{r}
denver<-subset(winner_year,winner_year$winner=="Denver Broncos")
max(denver$year)

```


# create 'winner2', a copy of the vector 'winner'
```{r}
winner2<-winner
winner2
```


# replace some team names in 'winner2' as follows:
# "New York Giants" to "NYG"
# "New York Jets" to "NYJ"
# "Kansas City Chiefs" to "KCC"
```{r}
winner2<-replace(winner2,which(winner2=="New York Giants"),"NYG")
winner2<-replace(winner2,which(winner2=="New York Jets"),"NYJ")
winner2<-replace(winner2,which(winner2=="Kansas City Chiefs"),"KCC")
winner2
```




# =====================================================
# Losing team
# Write the commands to answer the following questions
# =====================================================

# get the losing teams of the first 5 superbowls
```{r}
loser_year<-data.frame(loser,year)
loser_year<-order(loser_year,decreasing = FALSE)
head(loser_year)
```


# get the losing teams of the last 5 superbowls
```{r}
tail(loser_year)
```


# get the losing teams in even positions (2, 4, 6, 8, etc)
```{r}
loser[seq_along(loser) %%2 == 0]
```


# get the losing teams in odd positions (1, 3, 5, 7, etc)
```{r}
loser[seq_along(loser) %%2 > 0]
```


# create the frequency table 'los_freqs'
# of losing teams
```{r}
los_freqs=table(loser)
```


# What is the team that have lost the superbowl
# the most times?, and how many times?
```{r}
los_freqs<-sort(los_freqs,decreasing = TRUE)
los_freqs[1]
```




# =====================================================
# Winners and Losers
# Write the commands to answer the following questions
# =====================================================

# how many different teams have played the superbowl?
```{r}
teams<-append(winner,loser)
unique(teams)
```


# teams that have played the superbowl and have never lost
```{r}
unique(winner[!c(winner %in% loser)])
```


# teams that have played the superbowl and have never won
```{r}
unique(loser[!c(loser %in% winner)])
```


# teams that have played the superbowl (both won and lost)?
```{r}
won_lost<-unique(winner[c(winner %in% loser)])
won_lost
```


# how many teams have both won and lost the superbowl?
```{r}
print("Number of teams that have won and lost")
print(length(won_lost))
```



# what team won the superbowl in 2000
```{r}
for(i in 1:49)
{
if(winner_year$year[i]==2000)
{print(winner_year$winner[i])}}
```


# what team lost the superbowl in 2000
```{r}
loser[c(year %in% 2000)]
```


# What are the champions of years 1970, 1980, 1990, 2000 & 2010?
```{r}
for(i in 1:49)
{
if(winner_year$year[i]%%10==TRUE)
{print(winner_year$winner[i])}
}
```


# what teams won the superbowl in the 1970s (1970-1979)
```{r}
winner[c(year %in% c(1970:1979))]
```


# what teams lost the superbowl in the 1990s (1990-1999)
```{r}
loser[c(year %in% c(1990:1999))]
```


# create a data frame 'superbowl_df' with the three vectors:
# year, winner, loser
```{r}
superbowl_df<-data.frame(year,loser,winner)
superbowl_df
```


# create a list 'superbowl' with the three vectors:
# year, winner, loser (each vector as an element of the list)

```{r}
superbowl<-list(year,winner,loser)
superbowl
```

